Predicting detection probabilities for rare and undersampled North American landbirds

### 2023-10-20 article/article.tex
wrote discussion for removal modelling

* results somewhat underwhelming for cross validation, but both models showed some support that a multi species model can improve precision of estimates (removal model in particular)

*

* migration strategy mattering less

* talk about Common Redpoll removal modelling

>>>>

\par In this paper, we developed a multi-species model of detectability by expanding upon the QPAD methodology developed by \citet{solymos_calibrating_2013}.
We applied the multi-species QPAD removal model and multi-species QPAD distance model to the large dataset of point counts collected by the NA-POPS project \citep{edwards_point_2023}, and compared estimates of cue rate and EDR from a multi-species context to estimates of cue rate and EDR from the previous single species context.
Given the results of these comparisons, we sought to predict these components of detectability for seven species of North American landbird that are rare and undersampled.

\subsection{Removal Modelling}
\par Estimates of cue rate changed very little from species that were originally modelled by the single-species models in \citet{edwards_point_2023}, to the same ones modelled in the multi-species model here.
The 5-fold cross validation tended to show either no preference for one model or another when sample size was large, and a preference (albeit an extremely small preference) for single-species models when sample size was smaller.
When sample size was small, a multi-species model tended to produce more precise estimates of cue rate.
Given the extremely small preference for single-species models under a lower sample size, it would appear that the bias-variance tradeoff is likely a welcome tradeoff here, because the decrease in variance is met by only a very slight increase in bias.

\par There were seven species where the relative difference in cue rate between the single-species model and multi-species model was over 20\% (noting that this threshold was chosen arbitrarily).
These species (Willow Ptarmagin, Rusty Blackbird, Red-breasted Sapsucker, Common Redpoll, Calliope Hummingbird, Allen's Hummingbird, and Bohemian Waxwing), mostly had fairly small sample sizes ($\leq 515$ for all species except Red-breasted Sapsucker [$n = 1854$] and Common Redpoll [$n = 2009$]), had the lowest cue rate in the single-species models, and had among the lowest cue rates in the multi-species model.
For removal modelling, all those species except for Common Redpoll ended up having the multi-species model chosen as the more predictive model, indicating that sharing information among related species tended to improve their estimates of cue rate.
Common Redpoll was likely an exception to this because the original estimate of cue rate in NA-POPS was a wild underestimate (cue rate of TO DO), likely because of a convergence issue in the \texttt{detect} package.
It could be worth running Common Redpoll using the Bayesian single species model developed here to see if the estimate of cue rate is still extremely low, or if the vague but still somewhat informative prior helps to bring the estimated parameter into a more sensible space.
It could also be worth investigating the merit of this type of modelling to estimate cue rate for species where the cue rate is extremely low.
In our case, 41 out of the 79 species that had the lowest 25\% quantile of cue rate in the single-species model ended up having the multi-species model chosen as the preferred model.
TO DO Re-read results of Martin-Schwarze.

\par Although this was not a primary goal of our study, we found much less support than \citet{solymos_phylogeny_2018} for migration strategy having an effect on cue rate.
Both the estimate effects for resident species and migratory species were effectively 0, with very little of the density above or below 0.025 of 0.
One potential reason is that we had proportionally more resident species in our analysis (85 out of 316; 26.9\%) compared to \citet{solymos_phylogeny_2018} (23 out of 141; 16.3\%), and so the richer information provided by these additional species could have brought the effect down to 0.
Additionally, while \citet{solymos_phylogeny_2018} only regressed estimates of cue rate against the predictor variables, we actually estimated cue rate within the context of the same predictors, and so the phylogenetic effect could have just overwhelmed the estimates, making the migratory strategy meaningless.

\subsection{Distance Modelling}
\par Estimates of EDR also changed very little from species that were originally modelled by the single-species models in \citet{edwards_point_2023}, to the same ones modelled in the multi-species model here.
However, there was evidence for an average small downward shift in EDRs when modelled under a multi-species context.
Still, the 5-fold cross validation tended to show little to no preference for one more or another when the sample size was large, and a slight (but still very small) preference for the multi-species model when the sample size was smaller.
There was less evidence for the multi-species model being able to improve the precision of EDR estimates across sample sizes, indicating that a single-species model may mostly be sufficient for precise estimates.
Under this context, it would appear that a multi-species model may slightly bring down estimates of EDR for species with a smaller sample size.

\par Talk about species with a large relative difference

\par Talk about parameters

\subsection{Predictions of Cue Rate and EDR}
